---
title: "dataviz_101_workshop"
author: hannah yan han
output: html_document
---

In this session, we'll be using R's ggplot2 package and its building blocks (aesthetic mapping, faceting, etc) to explore Diamonds dataset

A few ways you can run a code snippet:

1. press the green triangle
2. mouse inside the grey area, press Run> run current chunk
3. cmd+shift+enter (mac) to run it.


```{r setup, include = FALSE}
#if you already installed ggplot2, u can ignore this
install.packages("ggplot2")

#load libraries
library(ggplot2)
```

```{r}
#load diamonds dataset
data(diamonds)
#description of the dataset
help(diamonds)
```

```{r}
#view the top 6 rows of the dataset
head(diamonds)
```


```{r}
#get summary description of dataset
#notice how the summary differs for numerical and categorical fields
summary(diamonds)
```

#how one (or more) variables are distributed
##histogram
```{r}
#histogram
ggplot(diamonds, aes(x=price)) + 
  geom_histogram(binwidth=200, color='white') 
```
```{r}
ggplot(diamonds, aes(x=price, fill=clarity)) + 
  geom_histogram(binwidth=500, color='white')
```

```{r}
#instead of stacking, let's facetting.
ggplot(diamonds, aes(x=price, fill=clarity)) + 
  geom_histogram(binwidth=500, color='white') + 
  facet_grid(clarity~.)
```

# Explore how two variables are related - scatterplot, boxplot, tile plot
##scatter plot
```{r}
#scatter plot in black and white
ggplot(diamonds, aes(x = carat, y = price)) + 
  geom_point()
```
```{r}
#add some colors
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) + 
  geom_point()
```

```{r}
#now write your own code, color by cut instead of clarity
```

```{r}
#make dots smaller
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) + 
  geom_point(size=0.2)
```

```{r}
#vary the size
ggplot(diamonds, aes(x = carat, y = price, color = clarity, size = cut)) + 
  geom_point()
```

```{r}
#add transparency
ggplot(diamonds, aes(x = carat, y = price, color = clarity, size = cut, alpha = 0.5)) + geom_point()
```

```{r}
#fit a linear regression line
ggplot(diamonds, aes(x=carat, y=price)) + 
  geom_point() + 
  stat_smooth(method='lm', se=F)
```

```{r}
#fit a smooth curve, which looks more suitable here
ggplot(diamonds, aes(x=carat, y=price)) + geom_point() + geom_smooth()
```
```{r}
#fit multiple smooth curve
ggplot(diamonds, aes(x = carat, y = price, color = clarity, alpha = 0.5)) + 
  geom_point() + geom_smooth(se = FALSE)
```
Feeling more adventurous? We can show only one fitted line
```{r}
#fit multiple smooth curve
ggplot(diamonds, aes(x = carat, y = price, color = clarity, alpha = 0.5)) + 
  geom_point() + 
  stat_smooth(aes(group=1))
```


```{r}
#facet
ggplot(diamonds, aes(x=carat, y=price, color=clarity, alpha=0.5)) + 
  geom_point() + 
  facet_wrap(~ cut) 
```

```{r}
#remove the grey background for a minimalistic look
ggplot(diamonds, aes(x=carat, y=price, color=clarity, alpha=0.5)) + 
  geom_point() + 
  facet_wrap(~ cut) + 
  theme_minimal()
```



#boxplot
```{r}
ggplot(diamonds, aes(x=color, y=price, fill=color, alpha=0.8)) + 
  geom_boxplot()
```

```{r}
ggplot(diamonds, aes(x=color, y=price, fill=color, alpha=0.8)) + 
  geom_violin()
```

```{r}
#faceted boxplot
ggplot(diamonds, aes(x = clarity, y = price, color = cut)) + 
  geom_boxplot() + 
  facet_wrap(~cut) +
  theme_minimal() 
```

```{r}
#add title
ggplot(diamonds, aes(x = clarity, y = price, color = cut)) + 
  geom_boxplot() + 
  facet_wrap(~cut) +
  theme_minimal() +
  ggtitle('diamond price by cut and clarity')
```


Well done!
If you want to learn more abut ggplot2, you can refer to this cheatsheet: https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf
This tutorial is made with <3 for WWCode Singapore Dataviz workshop (2017 Jun)


